---
title: "Source_Code"
author: "s18840"
date: "2024-01-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#loading packages needed
```{r}
library(tidyverse)
library(skimr)
library(janitor)
```
```{r}
library(lubridate)
```

```{r}
library(dplyr)
```

```{r}
library(stringr)
```


#importing the data sets
```{r}
Local_data<-read_csv("../data/london_local_data_2022.csv",col_types = c('cfTddddddd'))
```


```{r}
skim(Local_data)
```
```{r}
glimpse(Local_data)
```
```{r}
local_site<-read_csv("../data/london_local_sites.csv",col_types = c('fcddf'))
```

```{r}
skim(local_site)
```
```{r}
glimpse(local_site)
```

```{r}
view(local_site)
view(Local_data)
```


#to check the 36 site names and codes
```{r}
unique_sites<-local_site%>%
  distinct(code,site)
print(unique_sites)
```

```{r}
write_csv(unique_sites,"../Latex_file/unique_sites.csv",col_names = TRUE)
```


#to check gas types
```{r}

gas_types <- colnames(Local_data)[6:12]
gas_types
```

#to get the gasses and no of sites that they are recorded
```{r}
parameter_counts <- local_site %>%
  count(Parameter_name) %>%
  rename("No of Sites" = n)%>%
  rename(Pollutants=Parameter_name)

parameter_counts
```

```{r}
write_csv(parameter_counts,"../Latex_file/parameter_counts.csv",col_names = TRUE)
```

```{r}
str(Local_data)
```


#replacing NA values with a zero
```{r}
Local_data[is.na(Local_data)] <- 0
```

#seperating date column in to more simmpler way

```{r}
Local_data<-Local_data%>%
  mutate(time = as.factor(format(as.POSIXct(date),format ="%H:%M:%S")),date=as.Date(date))%>%
  relocate(time, .after = date)

```


```{r}
Local_data <- Local_data %>%
  mutate(month =factor(case_when(
    month(date)==1 ~ "January",
    month(date)==2 ~ "February",
    month(date)==3 ~ "March",
    month(date)==4 ~ "April",
    month(date)==5 ~ "May",
    month(date)==6 ~ "June",
    month(date)==7 ~ "July",
    month(date)==8 ~ "August",
    month(date)==9 ~ "September",
    month(date)==10 ~ "October",
    month(date)==11 ~ "November",
    month(date)==12 ~ "December"
  )),.after = date)

Local_data$month <- factor(Local_data$month,levels = c("January","February","March","April","May","June","July","August","September","October","November","December"))
```



```{r}
head(Local_data)
```


#year 2022 analysis by gas
```{r}

Total_count <-Local_data %>% select(nox,no2,no,pm10,o3,pm2_5,so2) %>%
  summarise_all(sum) %>% t() %>%
  as.data.frame() %>% 
  rownames_to_column(var = "substance") %>%
  rename(total = V1)
Total_count

```



```{r}
yearly_graph_1<-Total_count%>%
  filter(substance != "nox")%>%
  ggplot(aes(x = substance, y = total, fill = substance))+
  geom_bar(stat = "identity", position = "dodge")+
  labs(x = "Substance",
       y = "Total Emissions")+
  theme_minimal()+
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K"))

yearly_graph_1
```

```{r}
postscript(file = "../Latex_file/Graphs/yearly_graph_1.eps",width = 10, height = 8, horizontal = F)
yearly_graph_1
dev.off()
```




#monthly analysis
```{r}
Total_per_month <- Local_data %>% group_by(month)%>%
  summarise(Total = sum(no)+sum(no2)+sum(nox)+
              sum(pm10)+sum(o3)+sum(pm2_5)+sum(so2)) 
Total_per_month


#top 6 months
Total_per_month %>%
  arrange(desc(Total)) %>%
  slice_head(n=6)

#least
Total_per_month %>%
  arrange(Total) %>%
  slice_head()
```

```{r}
Total_per_month %>%
  ggplot(aes(x = month, y = Total)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Emissions per Month",
       x = "Month",
       y = "Total Emissions") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))
```


#monthly analysis by gas
```{r}
Total_per_month_by_gas <- Local_data %>%
  group_by(month) %>%
  summarise(
    nox = sum(nox),
    no2 = sum(no2),
    no = sum(no),
    pm10 = sum(pm10),
    o3 = sum(o3),
    pm2_5 = sum(pm2_5),
    so2 = sum(so2)
  )

Total_per_month_by_gas
  
```
```{r}
Total_per_month_by_gas_2<-Total_per_month_by_gas%>%
  gather(key = "Gas", value = "Total", -month) %>%
  ggplot(aes(x = month, y = Total, fill = Gas)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
       x = "Month",
       y = "Total Emissions") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))+
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K"))
  
Total_per_month_by_gas_2

postscript(file = "../Latex_file/Graphs/Total_per_month_by_gas_2.eps",width = 10, height = 8, horizontal = F)
Total_per_month_by_gas_2
dev.off()

```
```{r}

```

```{r}
Total_per_month_by_gas_3<-Total_per_month_by_gas %>%
  gather(key = "Gas", value = "Total", -month) %>%
  ggplot(aes(x = month, y = Total, color = Gas, group = Gas)) +
  geom_point() +
  geom_line() +
  labs(
       x = "Month",
       y = "Total Emissions") +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K"))+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))

Total_per_month_by_gas_3

postscript(file = "../Latex_file/Graphs/Total_per_month_by_gas_3.eps",width = 10, height = 8, horizontal = F)
Total_per_month_by_gas_3
dev.off()
```


```{r}
hour_wise_data <- Local_data %>% group_by(time)%>%
  summarise(Total = sum(no)+sum(no2)+sum(nox)+
              sum(pm10)+sum(o3)+sum(pm2_5)+sum(so2)) 
hour_wise_data


#top 10 hours
hour_wise_data %>%
  arrange(desc(Total)) %>%
  slice_head(n=10)

#least
hour_wise_data %>%
  arrange(Total) %>%
  slice_head(n=5)
```



#Filtering observations by sites wich collect data about emitting same gas


```{r}
Sites_emit_no2<-local_site%>%filter(Parameter_name=="Nitrogen dioxide")
Sites_emit_PM10<-local_site%>%filter(Parameter_name=="PM10 particulate matter (Hourly measured)")
Sites_emit_o3<-local_site%>%filter(Parameter_name=="Ozone")
Sites_emit_PM2.5<-local_site%>%filter(Parameter_name=="PM2.5 particulate matter (Hourly measured)")
Sites_emit_so2<-local_site%>%filter(Parameter_name=="Sulphur dioxide")
```
```{r}
no2_local_data<-inner_join(Local_data,Sites_emit_no2,by=c("site", "code"))
PM10_local_data<-inner_join(Local_data,Sites_emit_PM10,by=c("site", "code"))
o3_local_data<-inner_join(Local_data,Sites_emit_o3,by=c("site", "code"))
PM2.5_local_data<-inner_join(Local_data,Sites_emit_PM2.5,by=c("site", "code"))
So2_local_data<-inner_join(Local_data,Sites_emit_so2,by=c("site", "code"))

```


```{r}
No2_map_df<-no2_local_data%>%select(site,longitude,latitude,no2)
PM10_map_df<-PM10_local_data%>%select(site,longitude,latitude,pm10)
o3_map_df<-o3_local_data%>%select(site,longitude,latitude,o3)
PM2.5_map_df<-PM2.5_local_data%>%select(site,longitude,latitude,pm2_5)
So2_map_df<-So2_local_data%>%select(site,longitude,latitude,so2)
```


#NO2analysis by site

```{r}
no2_by_site <- no2_local_data %>% group_by(site)%>%
  summarise(Total = sum(no2)) 
no2_by_site
```

```{r}
no2_by_site_graph<-no2_by_site %>%
  mutate(City = str_trim(str_extract(site, "^[^-]+"))) %>%
  ggplot(aes(x = City, y = Total, fill = City)) +
  geom_bar(stat = "identity") +
  labs(
       x = "City",
       y = "Total Emissions") +
  theme_minimal()+
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K"))+
  coord_flip()

no2_by_site_graph

postscript(file = "../Latex_file/Graphs/no2_by_site_graph.eps",width = 10, height = 8, horizontal = F)
no2_by_site_graph
dev.off()
  
```


```{r}
no2_by_Westminister <- no2_local_data %>% 
   filter(str_detect(site, "^Westminster"))%>%
  group_by(site)%>%
  summarise(Total = sum(no2)) 
no2_by_Westminister
```

```{r}
no2_by_Westminister_graph<-no2_by_Westminister %>%
  ggplot(aes(x = site, y = Total, fill = site)) +
  geom_bar(stat = "identity") +
  labs(
       x = "Site",
       y = "Total Emissions") +
  theme_minimal()+
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K"))+
  
  scale_x_discrete(labels = function(x) str_wrap(x, width = 15))+
  coord_flip()

no2_by_Westminister_graph

postscript(file = "../Latex_file/Graphs/no2_by_Westminister_graph.eps",width = 10, height = 8, horizontal = F)
no2_by_Westminister_graph
dev.off()


```

```{r}
No2_map_df %>%
  ggplot(aes(x = longitude, y = latitude, color = no2)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Scatter Plot of no2 based on Longitude and Latitude",
       x = "Longitude",
       y = "Latitude",
       color = "no2") +
  theme_minimal()
```


```{r}
No2_map<-No2_map_df %>%
  mutate(City = str_trim(str_extract(site, "^[^-]+"))) %>%
  ggplot(aes(x = longitude, y = latitude, size = no2,color = City)) +
  geom_point() +
  
  
  scale_size_continuous(range = c(2, 10)) +  # Adjust the range of sizes as needed
  labs(
       x = "Longitude",
       y = "Latitude",
       color = "City",
       size = "no2") +
  theme_minimal()

No2_map

postscript(file = "../Latex_file/Graphs/No2_map.eps",width = 15, height = 20, horizontal = F)
No2_map
dev.off()
```



#PM10 analysis by site


```{r}
PM10_by_site <- PM10_local_data %>% group_by(site)%>%
  summarise(Total = sum(pm10)) 
PM10_by_site
```



```{r}
PM10_by_site_graph<-PM10_by_site%>%
  mutate(City = str_trim(str_extract(site, "^[^-]+"))) %>%
  ggplot(aes(x = City, y = Total, fill = City)) +
  geom_bar(stat = "identity") +
  labs(
       x = "City",
       y = "Total Emissions") +
  theme_minimal()+
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K"))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))+
  coord_flip()

PM10_by_site_graph

postscript(file = "../Latex_file/Graphs/PM10_by_site_graph.eps",width = 10, height = 8, horizontal = F)
PM10_by_site_graph
dev.off()

  
  
```


```{r}
PM10_by_Wandsworth<- PM10_local_data %>% 
   filter(str_detect(site, "^Wandsworth"))%>%
  group_by(site)%>%
  summarise(Total = sum(pm10)) 
PM10_by_Wandsworth
```

```{r}
PM10_by_Wandsworth_graph<-PM10_by_Wandsworth%>%
  ggplot(aes(x = site, y = Total, fill = site)) +
  geom_bar(stat = "identity") +
  labs(
       x = "Site",
       y = "Total Emissions") +
  theme_minimal()+
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K"))+
  
  scale_x_discrete(labels = function(x) str_wrap(x, width = 15))


PM10_by_Wandsworth_graph

postscript(file = "../Latex_file/Graphs/PM10_by_Wandsworth_graph.eps",width = 10, height = 8, horizontal = F)
PM10_by_Wandsworth_graph
dev.off()

```



```{r}
PM10_map<-PM10_map_df %>%
  mutate(City = str_trim(str_extract(site, "^[^-]+"))) %>%
  ggplot(aes(x = longitude, y = latitude, size = pm10,color = City)) +
  geom_point() +
  
  
  scale_size_continuous(range = c(2, 10)) +  # Adjust the range of sizes as needed
  labs(
       x = "Longitude",
       y = "Latitude",
       color = "City",
       size = "PM10 amunt") +
  theme_minimal()

PM10_map

postscript(file = "../Latex_file/Graphs/PM10_map.eps",width = 10, height = 8, horizontal = F)
PM10_map
dev.off()
```






#PM2.5 analysis by site


```{r}
PM2.5_by_site <- PM2.5_local_data %>% group_by(site)%>%
  summarise(Total = sum(pm2_5)) 
PM2.5_by_site

```

```{r}
PM2.5_by_site_graph<-PM2.5_by_site%>%
 
  ggplot(aes(x = site, y = Total, fill = site)) +
  geom_bar(stat = "identity") +
  labs(
       x = "Site",
       y = "Total Emissions") +
  theme_minimal()+
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K"))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))

PM2.5_by_site_graph

postscript(file = "../Latex_file/Graphs/PM2.5_by_site_graph.eps",width = 10, height = 8, horizontal = F)
PM2.5_by_site_graph
dev.off()

  
```





#O3 analysis by site

```{r}

O3_by_site <- o3_local_data %>% group_by(site)%>%
  summarise(Total = sum(o3)) 
O3_by_site

```
```{r}
O3_by_site_graph<-O3_by_site%>%
  mutate(short_site = str_trim(str_extract(site, "^[^-]+"))) %>%
  ggplot(aes(x = short_site, y = Total, fill = site)) +
  geom_bar(stat = "identity") +
  labs(
       x = "Site",
       y = "Total Emissions") +
  theme_minimal()+
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K"))

O3_by_site_graph

postscript(file = "../Latex_file/Graphs/O3_by_site_graph.eps",width = 10, height = 8, horizontal = F)
O3_by_site_graph
dev.off()
  
  
```


#so2 analysis by site

```{r}
so2_by_site <- So2_local_data %>% group_by(site)%>%
  summarise(Total = sum(so2)) 
so2_by_site
```

